---
title: 'Session 3, Toy Shop With Integer Variables'
author: "Ralph Asher, ralph@datadrivensupplychain.com"
date: "8/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ompr)
library(ompr.roi)
library(ROI.plugin.glpk)
rm(list=ls())

```

We'll use the final example from the first toyshop exercise, but make the decision variables integers.

```{r Toy Shop Using Dataframes}

toyshop_df <- cbind.data.frame(
  Toy = c("Car", "Train", "Truck", "Airplane"),
  Time_Required = c(5, 3, 4, 6),
  Wood_Required = c(1, 1, 1.5, 2),
  Profit = c(7, 4, 6, 9),
  stringsAsFactors=FALSE)

toy_type_count <- nrow(toyshop_df)



toymodel <- MIPModel() %>%
  add_variable(toys[i], i= 1:toy_type_count, type='integer', lb=0) %>% 
  add_constraint(sum_expr(toys[i] * toyshop_df$Time_Required[i], i=1:toy_type_count) <= 40) %>%
  add_constraint(sum_expr(toys[i] * toyshop_df$Wood_Required[i], i=1:toy_type_count) <= 10) %>%
  set_objective(sum_expr(toys[i] * toyshop_df$Profit[i], i=1:toy_type_count))


toymodel

my_solution <- solve_model(toymodel,with_ROI("glpk",verbose=TRUE))

my_solution

toys_solution <- get_solution(my_solution,toys[i]) %>%
  dplyr::mutate(Toy = toyshop_df$Toy[i])


toys_solution


```
When the decision variables were continuous, the optimal solution was:

5 cars, 0 trains, 0 trucks, and 2.5 airplanes
objective value: 5*7 + 0*4 + 0*6 + 2.5*9 = 57.5

How does this objective value compare?  How do the solutions compare?

Which is more useful- the continuous decision variable results, or the integer variable results?